Two Decades of Smalltalk VM Development

### 2018-08-17 paper.tex
Optional footnote for Lemming Debugging ;-)

Debugging graph rewriting garbage collectors can be hard; bugs leave the heap in a scrambled state, and figuring out what happened when one is left with a corrupted graph, often too broken to traverse and make sense of, or apply tools to, is often effectively impossible.  But because the simulation is simply an object graph it is easy to copy the entire simulation. Therefore the simulator has been extended to by default take a copy of itself on every scavenge or old space reclamation, and to run the collection in the copy, performing a leak check afterwards (if enabled).  Hence if a bug occurs or validation fails it will fail in the copy while the original remains as is, yet to be corrupted.  We then take a fresh copy and use it to debug, repeating the process as many times as we like, an approach we call Lemming Debugging.  One throws the copy off the cliff and if it gets to the bottom without incident it is descarded and the simulation continues in the original.  This approach was key in rapid development of the new memory manager that supports both 32 and 64 bits \cite{Forwarders}.

>>>>

Debugging graph rewriting garbage collectors can be hard; bugs leave the heap in a scrambled state, and figuring out what happened when one is left with a corrupted graph, often too broken to traverse and make sense of, or apply tools to, is often effectively impossible.  But because the simulation is simply an object graph it is easy to copy the entire simulation. Therefore the simulator has been extended to by default take a copy of itself on every scavenge or old space reclamation, and to run the collection in the copy, performing a leak check afterwards (if enabled).  Hence if a bug occurs or validation fails it will fail in the copy while the original remains as is, yet to be corrupted.  We then take a fresh copy and use it to debug, repeating the process as many times as we like, an approach we call Lemming Debugging \footnote{http://www.romainbrasier.fr/404/}.  One throws the copy off the cliff and if it gets to the bottom without incident it is descarded and the simulation continues in the original.  This approach was key in rapid development of the new memory manager that supports both 32 and 64 bits \cite{Forwarders}.
